Project for determing the class of a given exercise 
========================================================
Outline - I am downloading the data for the training set and I am partitioning it into two parts of sizes 75% and 25%. The former will be my training set and the later will be my testing set within the larger framework of the training set I will test my model on the testing part and check the quality of the predictions and once satisfied, it will be applied to the **testing set provided** to us. 

```{r}
#Downloading data 
fileUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
download.file(fileUrl, destfile="pml-training.csv", method="curl")
FitbitTrain<-read.csv("pml-training.csv")
```
We are creating the test and train sets
```{r}
library(caret);library(psych)
#important- here I am using stringsAsFactors=FALSE, so it will convert the classe variable into a character variable and not leave it as a factor variable as it is).
#Here an indexing set it being created using classe which will be the object we 
#want to predict later on. 
trainIndex <- createDataPartition(FitbitTrain$classe,p=0.75,list=FALSE)
traininginFitbitTrain<-FitbitTrain[trainIndex,]
testinginFitbitTrain<- FitbitTrain[-trainIndex,]
##dim(FitbitTrain); dim(trainIndex); were checked to be correct.
###Now I am trying to reduce the variables that I need to fit the data
##Step 1 is to determine the onces which enteries are characters,looking that data
## shows that the enteries which are characters, seem to be derived from the data or is 
##information like name, time, date, etc. 
#This puts the list of all such columns together
class <- c()
for (i in 1:160){
      if (class(FitbitTrain[,i])=="factor")
      {class <- c(class,i)}
}
class
##Now we remove variables with very small variance, for this we list those variables
zerovar <- nearZeroVar(FitbitTrain)
## Now we remove the enteries which are sparse. For this we use the describe the function 
#which is available using the psych package.The second entry gives the number of row enteries. 
#These numbers vary between 19622 and 4** 
sparse <- which(describe(FitbitTrain)[,2]<19622) 
invalid <- append(append(sparse,zerovar),class)
invalid <- unique(sort(invalid))
##Here is put together all the features that I would like to ignore.  
invalid
#invalid <- invalid[-length(invalid)]
#trainIndex <- createDataPartition(FitbitTrain$classe,p=0.75,list=FALSE)
traininginFitbitTrain1<-FitbitTrain[trainIndex,-invalid]
#str(traininginFitbitTrain1)
testinginFitbitTrain1<- FitbitTrain[-trainIndex,-invalid]
#str(testinginFitbitTrain1)

##dim(FitbitTrain); dim(trainIndex); were checked to be correct.
invalid1 <- invalid[1:length(invalid)-1]
invalid1
##I want to ensure that I put back the classe feature into my data:

traininginFitbitTrain2<-FitbitTrain[trainIndex,-invalid1]
testinginFitbitTrain2<-FitbitTrain[-trainIndex,-invalid1]


###Now I am doing the training and the train data


preProc <- preProcess(traininginFitbitTrain1, method="pca")
#str(traininginFitbitTrain1)
valuesintraininginFitbitTrain1 <-  predict(preProc,traininginFitbitTrain1)
#head(valuesintraininginFitbitTrain2)
#head(traininginFitbitTrain2)

modelintraininginFitbitTrain2 <- train(traininginFitbitTrain2$classe ~., data=valuesintraininginFitbitTrain1, method="rf", trControl = trainControl(method = "cv", number = 4, allowParallel = TRUE,verboseIter = TRUE))
#modelintraininginFitbitTrain2

#str(valuesintraininginFitbitTrain1)
valuesintestinginFitbitTrain1 <-predict(preProc,testinginFitbitTrain1)

confusionMatrix(testinginFitbitTrain2$classe, predict(modelintraininginFitbitTrain2,valuesintestinginFitbitTrain1))

#################Now predicting the outcome of the 20 cases given in the assignment

fileUrl <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
download.file(fileUrl, destfile="pml-testing.csv", method="curl")
FitbitTest<-read.csv("pml-testing.csv")
FitbitTest1<-FitbitTest[,-invalid]
valuesinFitbitTest1 <-predict(preProc,FitbitTest1)

answers <- predict(modelintraininginFitbitTrain2,valuesinFitbitTest1)
#predict(modelintraininginFitbitTrain2,valuesintestinginFitbitTrain1)
answers

```
